Note on: Out-of-Sample Error for Laplacian Eigenmaps

### 2021-02-23 notes/weekly_notes/12.24.20.NotesForWeek.tex
Updated the note on insights and connections.

The estimate we will consider is a kernel smoother passed over $\wh{\theta}$,
\begin{equation}
\label{eqn:kernel_smoother_laplacian_eigenmaps}
\wh{f}(x) := T_{\varepsilon,n}\wh{\theta}_{1:n}(x).
\end{equation}
In order to relate the out-of-sample error $\|\wh{f} - f^{\ast}\|_P$ to the in-sample error $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n$, we will assume the following.
\begin{enumerate}[label=(A\arabic*)]
	\item
	\label{asmp:kernel}
	The kernel function $\eta$ is supported on a subset of $[0,1]$, with $\eta(1/2) > 0$. Additionally, $\eta$ is \textcolor{red}{Lipschitz} on its support.
	\item 
	\label{asmp:domain} There exists $\alpha > 0$ such that for every $x \in \mc{X}$, $d_{\nu}(x) \geq \alpha \cdot \varepsilon^d$.
	\item 
	\label{asmp:density}The distribution $P$ admits a density $p$ with respect to Lebesgue measure. The density $p \in C^1(\mc{X})$, and is bounded away from $0$ and $\infty$ for all $x \in \mc{X}$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) < p_{\max} < \infty.
	\end{equation*}
\end{enumerate}
Under Assumption~\ref{asmp:kernel}, we automatically have that $d_\nu(x) = \varepsilon^d$ for all $x \in \mc{X}_{\varepsilon}$. Assumption~\ref{asmp:domain} is thus a condition on the boundary $\partial \Omega$, and is implied for all sufficiently small $\varepsilon$ if the boundary is Lipschitz. 
\begin{lemma}
	\label{lem:kernel_smoothing_laplacian_eigenmaps}
	Suppose $f^{\ast} \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$. Then with probability at least $1 - \delta - C\varepsilon^d\exp\{-Cn\varepsilon^d\}$, it holds that
	\begin{equation*}
	\|\wh{f} - f^{\ast}\|_P^2 \leq C\biggl(\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n^2 + \frac{1}{\delta} \cdot \frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2\biggr)
	\end{equation*}
\end{lemma}
Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data (on the order of $1/(n\varepsilon^d)$) by a factor of $\varepsilon^2$. The bias term, on the other hand, is standard, and in particular the following estimate of the bias can be obtained. 
\begin{lemma}
	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{\beta}(\Omega)$ for $\beta = s - 1$, and that $\eta$ is an order-$s$ kernel. Then
	\begin{equation*}
	\|T_{P,\varepsilon}f^{\ast} - f^{\ast}\|_P^2 \leq C \varepsilon^{2s}.
	\end{equation*}
\end{lemma}

Thus balancing bias and variance, using a smaller bandwidth ($\varepsilon \asymp n^{-1/(2\beta + d)}$) than is typical for kernel smoothing with noisy data, leads to an asymptotically negligible error ($n^{-2s/(2\beta + d)}$) compared to the minimax rate for estimating Sobolev functions.

@@

	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to constant) a contraction, meaning that for any $u \in L^2(X)$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_1}
	\|T_{\varepsilon,n}u\|_P^2 \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \|u\|_n^2.
	\end{equation}
	We begin with a pointwise estimate, which follows from Jensen's inequality:
	\begin{equation*}
	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} u_i^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{equation*}
	(Note that it is here that we have used the positivity of $\eta(\cdot)$.) Integrating with respect to $P$ and exchanging sum with integral yields an upper bound on the $L^2(P)$ norm of $T_{n,\varepsilon}u$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_2}
	\|T_{n,\varepsilon}u\|_P^2 \leq \sum_{i = 1}^{n} u_i^2 \Bigl(\int_{\Omega} \frac{\eta(\|X_i - x\|/\varepsilon)}{n d_n(x)} p(x) \,dx \Bigr)
	\end{equation}
	The degree estimate $d_n(x)$ is quite close to $p(x)$. As we show in Lemma~\ref{lem:measure_of_ball_estimate}, if $\varepsilon = \omega((\log n/n)^{1/d})$ then the empirical measure $d_n(x)$ concentrates around $d_P(x)$ uniformly over $x \in \Omega$, 
	\begin{equation*}
	\sup_{x \in \Omega}|d_n(x) - d_P(x)| = o_P(\varepsilon^d);
	\end{equation*}
	moreover, it follows from Lipschitz continuity of $p$ that for any $x \in \Omega$,
	\begin{equation*}
	d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) p(x') \,dx' = p(x) d_{\nu}(x) + O(\varepsilon^{d + 1});
	\end{equation*}
	and thus by the triangle inequality $\sup_{x \in \Omega}|d_n(x) - d_\nu(x) p(x)| = o_P(\varepsilon^d)$.

>>>>

The estimate we will consider is a kernel smoother passed over $\wh{\theta}$,
\begin{equation}
\label{eqn:kernel_smoother_laplacian_eigenmaps}
\wh{f}(x) := T_{\varepsilon,n}\wh{\theta}_{1:n}(x).
\end{equation}
In order to relate the out-of-sample error $\|\wh{f} - f^{\ast}\|_P$ to the in-sample error $\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n$, we will assume the following.
\begin{enumerate}[label=(A\arabic*)]
	\item
	\label{asmp:kernel}
	The kernel function $\eta$ is supported on a subset of $[0,1]$, with $\eta(1/2) > 0$. Additionally, $\eta$ is \textcolor{red}{Lipschitz} on its support.
	\item 
	\label{asmp:domain} There exists $\alpha > 0$ such that for every $x \in \mc{X}$, $d_{\nu}(x) \geq \alpha \cdot \varepsilon^d$.
	\item 
	\label{asmp:density}The distribution $P$ admits a density $p$ with respect to Lebesgue measure. The density $p \in C^1(\mc{X})$, and is bounded away from $0$ and $\infty$ for all $x \in \mc{X}$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) < p_{\max} < \infty.
	\end{equation*}
\end{enumerate}
Under Assumption~\ref{asmp:kernel}, we automatically have that $d_\nu(x) = \varepsilon^d$ for all $x \in \mc{X}_{\varepsilon}$. Assumption~\ref{asmp:domain} is thus a condition on the boundary $\partial \Omega$, and is implied for all sufficiently small $\varepsilon$ if the boundary is Lipschitz. 
\begin{lemma}
	\label{lem:kernel_smoothing_laplacian_eigenmaps}
	Suppose $f^{\ast} \in H^1(\mc{X})$ and $p \in C^1(\mc{X})$. Then with probability at least $1 - \delta - C\varepsilon^d\exp\{-Cn\varepsilon^d\}$, it holds that
	\begin{equation*}
	\|\wh{f} - f^{\ast}\|_P^2 \leq C\biggl(\|\wh{\theta}_{1:n} - f^{\ast}_{1:n}\|_n^2 + \frac{1}{\delta} \cdot \frac{\varepsilon^2}{n\varepsilon^d} |f|_{H^1(\mc{X})}^2 + \frac{1}{\delta}\|T_{\varepsilon,P}f^{\ast} - f^{\ast}\|_P^2\biggr)
	\end{equation*}
\end{lemma}
Notice that the variance term in the above is smaller than the typical variance term for kernel smoothing of noisy data (on the order of $1/(n\varepsilon^d)$) by a factor of $\varepsilon^2$. The bias term, on the other hand, is standard, and in particular the following estimate of the bias can be obtained. 
\begin{lemma}
	Suppose $f^{\ast} \in H_0^{s}(\Omega)$, $p \in C^{r}(\Omega)$ for $r = s - 1$, and that $\eta$ is an order-$s$ kernel. Then
	\begin{equation*}
	\|T_{P,\varepsilon}f^{\ast} - f^{\ast}\|_P^2 \leq C \varepsilon^{2s}.
	\end{equation*}
\end{lemma}
\begin{proof}
	\textcolor{red}{(TODO)}
\end{proof}

Thus balancing bias and variance, using a smaller bandwidth ($\varepsilon \asymp n^{-1/(2r + d)}$) than is typical for kernel smoothing with noisy data, leads to an asymptotically negligible error ($n^{-2s/(2r + d)}$) compared to the minimax rate for estimating Sobolev functions.

@@

	The left hand side of~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is equal to $\|T_{\varepsilon,n}(\wh{\theta}_{1:n} - f^{\ast}_{1:n})\|_P$, and~\eqref{pf:kernel_smoothing_laplacian_eigenmaps_0} is implied by showing that the operator $T_{\varepsilon,n}: L^2(X) \to L^2(\Omega)$ is (up to constant) a contraction, meaning that for any $u \in L^2(X)$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_1}
	\|T_{\varepsilon,n}u\|_P^2 \leq \frac{1}{\alpha}\bigl(1 + o_P(1)\bigr) \|u\|_n^2.
	\end{equation}
	We begin with a pointwise estimate, which follows from Jensen's inequality:
	\begin{equation*}
	\Bigl(\bigl(T_{n,\varepsilon}u\bigr)(x)\Bigr)^2 \leq \frac{1}{d_n(x) n} \sum_{i = 1}^{n} (u_i)^2 \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr).
	\end{equation*}
	(Note that it is here that we have used the positivity of $\eta(\cdot)$.) Integrating with respect to $P$ and exchanging sum with integral yields an upper bound on the $L^2(P)$ norm of $T_{n,\varepsilon}u$,
	\begin{equation}
	\label{pf:kernel_smoothing_laplacian_eigenmaps_2}
	\|T_{n,\varepsilon}u\|_P^2 \leq \sum_{i = 1}^{n} u_i^2 \Bigl(\int_{\Omega} \frac{\eta(\|X_i - x\|/\varepsilon)}{n d_n(x)} p(x) \,dx \Bigr)
	\end{equation}
	The degree estimate $d_n(x)$ is quite close to $p(x)$. As we show in Lemma~\ref{lem:measure_of_ball_estimate}, if $\varepsilon = \omega((\log n/n)^{1/d})$ then the empirical measure $d_n(x)$ concentrates around $d_P(x)$ uniformly over $x \in \Omega$, 
	\begin{equation*}
	\sup_{x \in \Omega}|d_n(x) - d_P(x)| = o_P(\varepsilon^d);
	\end{equation*}
	moreover, it follows from Lipschitz continuity of $p$ that for any $x \in \Omega$,
	\begin{equation*}
	d_P(x) = \int_{\Omega} \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) p(x') \,dx' = p(x) d_{\nu}(x) + O(\varepsilon^{d + 1});
	\end{equation*}
	and thus by the triangle inequality $\sup_{x \in \Omega}|d_n(x) - d_\nu(x) p(x)| = o_P(\varepsilon^d)$.
